Analysis of cloud enhancement events in a 30-year record of global solar irradiance at Thessaloniki, Greece

### 2024-09-09 article/article.Rmd
Commit 2024-09-09 11:11

The duration of the CE events is variable and can last for several minutes or even
more than an hour. To study the characteristics of these consecutive events, we
grouped them into bins of increasing duration in steps of one minute.  We have
identified `r ST_G0[, .N]` groups of CE events in the whole period of study, where
the group with the longest duration of
`r ST_G0[, max(GLB_diff.N)]` minutes occurred on
`r strftime(ST_G0[which.max(GLB_diff.N), Date], "%d\\ %B %Y")`
with SZA ranging between
$`r round(ST_G0[which.max(GLB_diff.N), SZA.min],1)`^\circ$ and
$`r round(ST_G0[which.max(GLB_diff.N), SZA.max],1)`^\circ$.
Figure\nobreakspace{}\ref{fig:ceg-duration-distribution} shows the frequency
distribution of the CE events according to their duration.  We conclude that although
some groups of events last for more than an hour, about $80\,\%$ have a duration of
less than 5 minutes.

($`r ceiling(100*ST_G0[,sum(GLB_ench.N>5)/.N])`\,\%$)

@@

The relation between the duration and the mean OI of the groups has also been studied
(Figure\nobreakspace{}\ref{fig:group-2d}).  Evidently, events of high excess
irradiation have small duration and vice versa. The vast majority of grouped events
are associated with small excess irradiation ($<5\,\text{kJ}/\text{m}^2$) and small
duration ($<5\,\text{min}$). For example the extra energy flux of
$5\,\text{kJ}/\text{m}^2$ for $5\,\text{min}$ is equivalent to OI of
$\sim17\,\text{W}/\text{m}^2$ for the same duration. While groups with strong excess
irradiation and long duration are very rare.  Similar results of this relation have
been reported by @Zhang2018 in a study using a far higher sampling rate
($100\,\text{Hz}$) than
ours.

```{r group-2d, echo=FALSE, fig.cap="Relation of excess irradiation of CE groups with their duration for Thessaloniki in the period 1994 -- 2023. The logarithmic color scale denotes the frequency of the respective groups of events."}
knitr::include_graphics("../images/P-groups-bin2d-1.png")
```

@@

An aspect of the CE events that is commonly reported and has some significance on the
solar energy production infrastructure are the extreme CE events (ECE), where solar
irradiance exceeds the expected irradiance on top of the atmosphere at the same SZA
(Equation\nobreakspace{}\ref{eq:ECE}).  Analogous to
Figure\nobreakspace{}\ref{fig:relative-month-occurrences} we have computed the
distribution of the number of occurrences of ECE events by month in
Figure\nobreakspace{}\ref{fig:relative-month-occurancies-ECE}. The most active period
for ECE events is in spring and early summer (March -- June), followed by a
period in early autumn (September and October).  This is probably related to the
weather characteristics in these periods, with frequent alternations between cloud-free periods and clouds.

```{r relative-month-occurancies-ECE, echo=FALSE, fig.cap="Seasonal variability of the number of ECE events relative to the the corresponding number of GHI measurements in Thessaloniki for the period 1994 -- 2023, in the form of a box and whisker plot. The box contains the data between the lower $25\\,\\%$ and the upper $75\\,\\%$ percentiles. The thick horizontal line and the diamond symbol represent the median and the mean values, respectively. The vertical lines (whiskers) extend between the maximum and minimum monthly values."}
knitr::include_graphics("../images/P-ECE-climatology-normlz-1.png")
```

<!--
```{r relative-month-occurancies-ECEs, echo=FALSE, fig.cap="Seasonal variability of the number of ECE events in Thessaloniki for the period 1994 -- 2023 normalized to the maximum occurring in March, in the form of a box and whisker plot. The box contains the data between the lower $25\\,\\%$ and the upper $75\\,\\%$ percentiles. The thick horizontal line and the diamond symbol represent the median and the mean values, respectively. The vertical lines (whiskers) extend between the maximum and minimum monthly values."}
knitr::include_graphics("../images/clim-ECE-month-norm-MAX-median-N-2.png")
```
-->

The distribution of the ECE events
(Figure\nobreakspace{}\ref{fig:P-extreme-distribution}) shows that in rare cases the
OI can exceed the TSI at TOA for the same SZA even by more than
$400\,\text{W}/\text{m}^2$, while in $75\,\%$ of the cases the OI is below
$200\,\text{W}/\text{m}^2$. The OI for the most frequent ECEs ranges between $140$
and $180\,\text{W}/\text{m}^2$.  These findings are in accordance with the results of
@Vamvakas2020, the only difference being that the OI values reported for are higher
than those for Thessaloniki.

```{r P-extreme-distribution, echo=FALSE, fig.cap="Distribution of ECE events in Thessaloniki for the period 1994 -- 2023. Measured as extra irradiance over TSI at TOA for the same SZA."}
knitr::include_graphics("../images/P-extreme-distribution-2.png")
```

>>>>

The duration of the CE events is variable and can last for several minutes or even
more than an hour. To study the characteristics of these consecutive events, we
grouped them into bins of increasing duration in steps of one minute.  We have
identified `r ST_G0[, .N]` groups of CE events in the whole period of study, where
the group with the longest duration of
`r ST_G0[, max(GLB_diff.N)]` minutes occurred on
`r strftime(ST_G0[which.max(GLB_diff.N), Date], "%d\\ %B %Y")`
with SZA ranging between
$`r round(ST_G0[which.max(GLB_diff.N), SZA.min],1)`^\circ$ and
$`r round(ST_G0[which.max(GLB_diff.N), SZA.max],1)`^\circ$.
Figure\nobreakspace{}\ref{fig:ceg-duration-distribution} shows the frequency
distribution of the CE events according to their duration.  We conclude that although
some groups of events last for more than an hour, about
$`r round(100*ST_G0[,sum(GLB_ench.N<=5)/.N],0)`\,\%$ have a duration of
equal or less than 5 minutes.

@@

The relation between the duration and the mean OI of the CE groups is shown in
Figure\nobreakspace{}\ref{fig:group-2d}.  As expected, events with small duration and
low OI are accumulated close to the origin of the plot.  Events of high excess
irradiation have a short duration and vice versa. The vast majority of grouped events
are associated with small excess irradiation ($<5\,\text{kJ}/\text{m}^2$) and short
duration ($<5\,\text{min}$). For example, the excess irradiation of
$5\,\text{kJ}/\text{m}^2$ for $5\,\text{min}$ is equivalent to an OI of
$16.7\,\text{W}/\text{m}^2$ with duration of $5\,\text{min}$. On the contrary, groups
with strong excess irradiation and long duration are very rare.  Similar results of
this relation have been reported by @Zhang2018 in a study using a far higher sampling
rate ($100\,\text{Hz}$) than ours.

```{r group-2d, echo=FALSE, fig.cap="Relation of excess irradiation of CE groups with their duration for Thessaloniki in the period 1994 -- 2023. The logarithmic color scale denotes the frequency of the respective groups."}
knitr::include_graphics("../images/P-groups-bin2d-1.png")
```

@@

An aspect of the CE events that is commonly reported and has some significance on the
solar energy production infrastructure are the extreme CE (ECE) events, where the
measured GHI exceeds the expected irradiance at the TOA at the same SZA
(Equation\nobreakspace{}\ref{eq:ECE}).  In analogy to
Figure\nobreakspace{}\ref{fig:relative-month-occurrences}, we show in 
Figure\nobreakspace{}\ref{fig:relative-month-occurancies-ECE} the number of ECE
events relative to the total number of one-min observations in each month. Judging by
the distribution of the monthly medians, the most active months in ECE events are
January -- March and November and the least active are July and August. The highest
interannual variability in the 30-year record occurs from October to April. In the
summer period, the fraction of ECE events is very small, with almost zero variability
in July and August.  Although the most active months for CE events are May and June
(Figure\nobreakspace{}\ref{fig:relative-month-occurrences}), these months behave
differently for the ECE events, suggesting that the cumulus clouds that usually cause
enhancements of irradiance are incapable to produce extreme events, possibly because
of their size and structure, in conjunction with the smaller SZAs in the summer.

In general, the fraction of ECE events during the year is very small, well below
$0.1\,\%$ of the data, therefore on average it would not affect the production of
solar energy. However, during short, isolated periods extreme enhancements of GHI can
be of concern.

```{r relative-month-occurancies-ECE, echo=FALSE, fig.cap="Seasonal variability of the relative occurrence of ECE events in Thessaloniki for the period 1994 -- 2023, in the form of a box and whisker plot. The box contains the data between the lower $25\\,\\%$ and the upper $75\\,\\%$ percentiles, with the thick horizontal line representing the median. The vertical lines (whiskers) extend between the maximum and minimum monthly values."}
knitr::include_graphics("../images/P-ECE-climatology-normlz-1.png")
```
<!--
```{r relative-month-occurancies-ECEs, echo=FALSE, fig.cap="Seasonal variability of the number of ECE events in Thessaloniki for the period 1994 -- 2023 normalized to the maximum occurring in March, in the form of a box and whisker plot. The box contains the data between the lower $25\\,\\%$ and the upper $75\\,\\%$ percentiles. The thick horizontal line and the diamond symbol represent the median and the mean values, respectively. The vertical lines (whiskers) extend between the maximum and minimum monthly values."}
knitr::include_graphics("../images/clim-ECE-month-norm-MAX-median-N-2.png")
```
-->

The distribution of the ECE events
(Figure\nobreakspace{}\ref{fig:P-extreme-distribution}) shows that in rare cases the
measured GHI exceeds the TOA TSI at the same SZA even by more than
$200\,\text{W}/\text{m}^2$, while in $75\,\%$ of the cases the OI is below
$40\,\text{W}/\text{m}^2$. The $50\,\%$ OI of the most frequent ECEs cases ranges
between $127$ and $211\,\text{W}/\text{m}^2$.  These findings are in accordance with
the results of @Vamvakas2020, the only difference being that the OI values reported
are higher than those for Thessaloniki.

```{r P-extreme-distribution, echo=FALSE, fig.cap="Relative frequency distribution of ECE events, in Thessaloniki for the period 1994 -- 2023, as a function of the extra irradiance over TSI at TOA for the same SZA."}
knitr::include_graphics("../images/P-extreme-distribution-2.png")
```

### 2024-09-09 article/article.tex
Commit 2024-09-09 11:11

The duration of the CE events is variable and can last for several
minutes or even more than an hour. To study the characteristics of these
consecutive events, we grouped them into bins of increasing duration in
steps of one minute. We have identified 34849 groups of CE events in the
whole period of study, where the group with the longest duration of 140
minutes occurred on 07~July 2013 with SZA ranging between \(52.1^\circ\)
and \(77.9^\circ\).
Figure\nobreakspace{}\ref{fig:ceg-duration-distribution} shows the
frequency distribution of the CE events according to their duration. We
conclude that although some groups of events last for more than an hour,
about \(80\,\%\) have a duration of less than 5 minutes.

(\(21\,\%\))

@@

The relation between the duration and the mean OI of the groups has also
been studied (Figure\nobreakspace{}\ref{fig:group-2d}). Evidently,
events of high excess irradiation have small duration and vice versa.
The vast majority of grouped events are associated with small excess
irradiation (\(<5\,\text{kJ}/\text{m}^2\)) and small duration
(\(<5\,\text{min}\)). For example the extra energy flux of
\(5\,\text{kJ}/\text{m}^2\) for \(5\,\text{min}\) is equivalent to OI of
\(\sim17\,\text{W}/\text{m}^2\) for the same duration. While groups with
strong excess irradiation and long duration are very rare. Similar
results of this relation have been reported by \citet{Zhang2018} in a
study using a far higher sampling rate (\(100\,\text{Hz}\)) than ours.

@@

\caption{Relation of excess irradiation of CE groups with their duration for Thessaloniki in the period 1994 -- 2023. The logarithmic color scale denotes the frequency of the respective groups of events.}\label{fig:group-2d}
\end{figure}

@@

An aspect of the CE events that is commonly reported and has some
significance on the solar energy production infrastructure are the
extreme CE events (ECE), where solar irradiance exceeds the expected
irradiance on top of the atmosphere at the same SZA
(Equation\nobreakspace{}\ref{eq:ECE}). Analogous to
Figure\nobreakspace{}\ref{fig:relative-month-occurrences} we have
computed the distribution of the number of occurrences of ECE events by
month in Figure\nobreakspace{}\ref{fig:relative-month-occurancies-ECE}.
The most active period for ECE events is in spring and early summer
(March -- June), followed by a period in early autumn (September and
October). This is probably related to the weather characteristics in
these periods, with frequent alternations between cloud-free periods and
clouds.

@@

\caption{Seasonal variability of the number of ECE events relative to the the corresponding number of GHI measurements in Thessaloniki for the period 1994 -- 2023, in the form of a box and whisker plot. The box contains the data between the lower $25\,\%$ and the upper $75\,\%$ percentiles. The thick horizontal line and the diamond symbol represent the median and the mean values, respectively. The vertical lines (whiskers) extend between the maximum and minimum monthly values.}\label{fig:relative-month-occurancies-ECE}
\end{figure}

The distribution of the ECE events
(Figure\nobreakspace{}\ref{fig:P-extreme-distribution}) shows that in
rare cases the OI can exceed the TSI at TOA for the same SZA even by
more than \(400\,\text{W}/\text{m}^2\), while in \(75\,\%\) of the cases
the OI is below \(200\,\text{W}/\text{m}^2\). The OI for the most
frequent ECEs ranges between \(140\) and \(180\,\text{W}/\text{m}^2\).
These findings are in accordance with the results of
\citet{Vamvakas2020}, the only difference being that the OI values
reported for are higher than those for Thessaloniki.

@@

\caption{Distribution of ECE events in Thessaloniki for the period 1994 -- 2023. Measured as extra irradiance over TSI at TOA for the same SZA.}\label{fig:P-extreme-distribution}
\end{figure}

>>>>

The duration of the CE events is variable and can last for several
minutes or even more than an hour. To study the characteristics of these
consecutive events, we grouped them into bins of increasing duration in
steps of one minute. We have identified 34849 groups of CE events in the
whole period of study, where the group with the longest duration of 140
minutes occurred on 07~July 2013 with SZA ranging between \(52.1^\circ\)
and \(77.9^\circ\).
Figure\nobreakspace{}\ref{fig:ceg-duration-distribution} shows the
frequency distribution of the CE events according to their duration. We
conclude that although some groups of events last for more than an hour,
about \(79\,\%\) have a duration of equal or less than 5 minutes.

@@

The relation between the duration and the mean OI of the CE groups is
shown in Figure\nobreakspace{}\ref{fig:group-2d}. As expected, events
with small duration and low OI are accumulated close to the origin of
the plot. Events of high excess irradiation have a short duration and
vice versa. The vast majority of grouped events are associated with
small excess irradiation (\(<5\,\text{kJ}/\text{m}^2\)) and short
duration (\(<5\,\text{min}\)). For example, the excess irradiation of
\(5\,\text{kJ}/\text{m}^2\) for \(5\,\text{min}\) is equivalent to an OI
of \(16.7\,\text{W}/\text{m}^2\) with duration of \(5\,\text{min}\). On
the contrary, groups with strong excess irradiation and long duration
are very rare. Similar results of this relation have been reported by
\citet{Zhang2018} in a study using a far higher sampling rate
(\(100\,\text{Hz}\)) than ours.

@@

\caption{Relation of excess irradiation of CE groups with their duration for Thessaloniki in the period 1994 -- 2023. The logarithmic color scale denotes the frequency of the respective groups.}\label{fig:group-2d}
\end{figure}

@@

An aspect of the CE events that is commonly reported and has some
significance on the solar energy production infrastructure are the
extreme CE (ECE) events, where the measured GHI exceeds the expected
irradiance at the TOA at the same SZA
(Equation\nobreakspace{}\ref{eq:ECE}). In analogy to
Figure\nobreakspace{}\ref{fig:relative-month-occurrences}, we show in
Figure\nobreakspace{}\ref{fig:relative-month-occurancies-ECE} the number
of ECE events relative to the total number of one-min observations in
each month. Judging by the distribution of the monthly medians, the most
active months in ECE events are January -- March and November and the
least active are July and August. The highest interannual variability in
the 30-year record occurs from October to April. In the summer period,
the fraction of ECE events is very small, with almost zero variability
in July and August. Although the most active months for CE events are
May and June
(Figure\nobreakspace{}\ref{fig:relative-month-occurrences}), these
months behave differently for the ECE events, suggesting that the
cumulus clouds that usually cause enhancements of irradiance are
incapable to produce extreme events, possibly because of their size and
structure, in conjunction with the smaller SZAs in the summer.

In general, the fraction of ECE events during the year is very small,
well below \(0.1\,\%\) of the data, therefore on average it would not
affect the production of solar energy. However, during short, isolated
periods extreme enhancements of GHI can be of concern.

@@

\caption{Seasonal variability of the relative occurrence of ECE events in Thessaloniki for the period 1994 -- 2023, in the form of a box and whisker plot. The box contains the data between the lower $25\,\%$ and the upper $75\,\%$ percentiles, with the thick horizontal line representing the median. The vertical lines (whiskers) extend between the maximum and minimum monthly values.}\label{fig:relative-month-occurancies-ECE}
\end{figure}

The distribution of the ECE events
(Figure\nobreakspace{}\ref{fig:P-extreme-distribution}) shows that in
rare cases the measured GHI exceeds the TOA TSI at the same SZA even by
more than \(200\,\text{W}/\text{m}^2\), while in \(75\,\%\) of the cases
the OI is below \(40\,\text{W}/\text{m}^2\). The \(50\,\%\) OI of the
most frequent ECEs cases ranges between \(127\) and
\(211\,\text{W}/\text{m}^2\). These findings are in accordance with the
results of \citet{Vamvakas2020}, the only difference being that the OI
values reported are higher than those for Thessaloniki.

@@

\caption{Relative frequency distribution of ECE events, in Thessaloniki for the period 1994 -- 2023, as a function of the extra irradiance over TSI at TOA for the same SZA.}\label{fig:P-extreme-distribution}
\end{figure}
